---
title: "Analysis.Rmd"
author: "Festus Yaboah, Chris Havenstein, and Cam Lu"
date: "July 7, 2017"
output: github_document
---

### Please set your working directory to the Analysis folder.


Depending on the directory you cloned the "MSDS6306-CaseStudy1" into, the path to the "Analysis" folder should be similar to:

* ./root/CaseStudy1/Analysis

By modifying "Analysis.Rmd" in the Analysis folder, I set an absolute path to the Analysis folder on my computer.


```{r setwd}
setwd("C:/Users/Chris/Desktop/GIT home/root/CaseStudy1/Analysis")
```


Source Makefile.R

```{r makefile}
source("./Data/Makefile.R")

```



## Question 1 on Merged Data: 

### Merge the data based on country shortcode. How many of the IDs match?

```{r CountRows}
nrow(MergeData)
```

#### 187 country shortcodes match.



## Question 2 on Merged Data:

### Sort the data frame in ascending order by GDP (so United States is last).
### What is the 13th country in the resulting data frame?

```{r}
MergeData <- MergeData[order(MergeData$GDP),]

MergeData[13,3]
```

#### Answer: The 13th Country in the reordered data is Samoa.
